---
title: "R Notebook"
output: html_notebook
---


## Background
- a little about destructive forces/ soil health --> I can write this later
--> for a site comparison= use all 3 

Addressing Question 1: Are destructive forces associated with decreasing microbial diversity? 
- What kind of disturbances impact beta diversity the greatest?


### Preprocessing

Loading in packages 

```{r}
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(ape)
library(ggtree)
library(vegan)
library(cowplot)
library(tidyverse)
library(data.table)
```

Loading in bacteria OTU 

```{r}
bac_10 = read.csv("../Data/Dryad_Microbial_OTU_Tables_10_Bacteria.csv")
bac_20 = read.csv("../Data/Dryad_Microbial_OTU_Tables_20_Bacteria.csv")
```

```{r}
temp <- list.files("../Data",pattern="Dryad_Microbial_OTU+.*csv", full.names = T) #creates an object that contains a string of file names with directory so they can be located

# Initializing list
Dat <- rep(0, length(temp)) # creates an empty vector used in the for loop
```

```{r}

for (i in seq_along(temp)) { # creates a list (I dont know if this is the right term) of dataframes
  Dat[[i]] <- read.csv(
    file = temp[[i]]
  )
}
```


```{r}
Dat
```


```{r}
for (i in seq_along(temp)) { 
  print(temp[i])
  data.frame(assign(sub(".csv", "",sub("../Data/Dryad_Microbial_OTU_Table*", "", temp[i])), data.frame(Dat[i])))
  
}
```

Making all column names upper case
```{r}
upper <- function(df) {
for( i in colnames(df)){
    colnames(df)[which(colnames(df)==i)] = toupper(i)
}
    return (df)
  
}

```

```{r}
s_10_Archaea = upper(s_10_Archaea)
s_20_Archaea = upper(s_20_Archaea)
s_10_Bacteria = upper(s_10_Bacteria)
s_20_Bacteria = upper(s_20_Bacteria)
s_10_Fungi = upper(s_10_Fungi)
s_20_Fungi = upper(s_20_Fungi)
```


We'll make a columnn that holds the entire phylogeny at the Family
** NEEDS TO BE GENERALIZED

```{r}
s_10_Archaea$Phylogeny <- paste0(s_10_Archaea$PHYLA, s_10_Archaea$CLASS, s_10_Archaea$ORDER, s_10_Archaea$FAMILY, sep = "")
s_20_Archaea$Phylogeny <- paste0(s_20_Archaea$PHYLA, s_20_Archaea$CLASS, s_20_Archaea$ORDER, s_20_Archaea$FAMILY, sep = "")

s_10_Bacteria$Phylogeny <- paste0(s_10_Bacteria$PHYLA, s_10_Bacteria$CLASS, s_10_Bacteria$ORDER, s_10_Bacteria$FAMILY, sep = "")
s_20_Bacteria$Phylogeny <- paste0(s_20_Bacteria$PHYLA, s_20_Bacteria$CLASS, s_20_Bacteria$ORDER, s_20_Bacteria$FAMILY, sep = "")

s_10_Fungi$Phylogeny <- paste0(s_10_Fungi$PHYLA, s_10_Fungi$CLASS, s_10_Fungi$ORDER, s_10_Fungi$FAMILY, sep = "")
s_20_Fungi$Phylogeny <- paste0(s_20_Fungi$PHYLA, s_20_Fungi$CLASS, s_20_Fungi$ORDER, s_20_Fungi$FAMILY, sep = "")
```


*** nnot working,.,???
```{r}
dfs = c(s_10_Bacteria)

for (i in seq_along(dfs)) { 
  dfs[[i]]$Phylogeny <- paste(dfs[[i]]$Phyla, dfs[[i]]$Class,  sep = "")
}
```




Making a function that removes the X prior to sample names and adding the "depth" of the sample

```{r}
changenames <- function(df, depth) {
for ( col in 1:ncol(df)){
    colnames(df)[col] <-  sub("X", depth, colnames(df)[col])
    
}
  return (df)
}
```

Applying the functions, overwrites s_10_Archaea
** STOPPED HERE

```{r}
s_10_Archaea = changenames(s_10_Archaea, 'depth10_')
#bac_20 = changenames(bac_20, 'depth20_')
```

In cases where the Phylogeny is the same across multiple rows, we will sum them

```{r}
bac_10_agglom = bac_10 %>%
  group_by(Phylogeny) %>%
  summarise_if(is.numeric, sum) %>% 
  select (-cluster)

bac_20_agglom = bac_20 %>%
  group_by(Phylogeny) %>%
  summarise_if(is.numeric, sum) %>% 
  select (-Cluster)
```

Checking the dimensions to see how different the number of phylogenies are
- the 0-10cm depth OTU table had more unique phylognies compareed to the 20-30cm OTU table
- Both have the same number of samples

```{r}
dim(bac_10_agglom)
dim(bac_20_agglom)

# find the number of unique phylogenies, we should expect 419 upon merging
length(unique(c(bac_10_agglom$Phylogeny, bac_20_agglom$Phylogeny)))
```

We're going to transpose the tables now--> need the column names

```{r}
bac_10_agglom_t =transpose(setDT(bac_10_agglom), make.names = 'Phylogeny', keep.names = 'Phylogeny')
bac_20_agglom_t =transpose(setDT(bac_20_agglom), make.names = 'Phylogeny', keep.names = 'Phylogeny')
```

Merging 

```{r}
bac = rbind(bac_10_agglom_t, bac_20_agglom_t,fill=TRUE)
```

Confirming the correct nunmber of columns! yes, we had 419 unique phylogennies and we created a bound datafraame with 419 unique phylogenies

```{r}
# We subtracct 1 because the first column holds the sample ids 
ncol(bac) -1
```


Adding a columnn that holds depth

```{r}
bac = bac %>% 
  mutate(depth = ifelse(grepl("depth10_", Phylogeny), '10',"20")) 
```

making the df to hold info
** ADD ENVIRONMENTAL DATA!!! USING EN FILE

```{r}
df = bac[, c("Phylogeny", "depth")]
```


Adding the environment table--> make column that matches depth10_39198 and merge

```{r}
en = read.csv("../Data/en.csv")

en = en %>% 
  mutate(Phylogeny = ifelse(grepl("010cm", Depth), paste0('depth10_',Sample), paste0('depth20_',Sample))) 
```

```{r}
describe = merge(df, en, by = 'Phylogeny')
```



Making the row names, the sample names so we don't lose them

```{r}
bac = bac %>% 
  remove_rownames %>% 
  column_to_rownames(var="Phylogeny")
```

Conversion to binary

```{r}
# need to drop the last column!!!!!
bac_bin<-subset(bac, select = -depth)
bac_bin[bac_bin>0]<-1
```

Calculating the pairwise distance of this binary matrix

```{r}
bac_bin_dist<-dist(bac_bin,method='binary')
```


```{r}
OTU_tree<-nj(bac_bin_dist)

ggtree(OTU_tree,layout="rectangular")%<+% 
  df + 
  geom_tiplab(aes(colour=depth))
```

Convert NA values to 0s
```{r}
bac_nona = bac %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), 0, .)))

```



also calcuulate the bray curtis...
```{r}
OTU_dist<- vegdist(subset(bac_nona, select = -depth),method="bray",binary=F) 
OTU_tree2<-nj(OTU_dist)
```





```{r}
ggtree(OTU_tree2) %<+% 
  df + 
  geom_tippoint(aes(color=depth)) 

```


Calculating NMDS

```{r}
set.seed(13)
NMDSdat<-metaMDS(OTU_dist,k=2) # k = 2 dimensions
```


```{r}
PDat<-data.frame(NMDS1=NMDSdat$points[,1],
                 NMDS2=NMDSdat$points[,2],
                 SampleID=row.names(bac))

PDat$depth = describe$depth

PDat$Clearcut = describe$Clearcut
PDat$Salvage = describe$Salvage
PDat$Slope = describe$Slope

qplot(x=NMDS1,y=NMDS2,colour=(Slope),alpha=I(0.6),data=PDat)+theme_bw()
```




