---
title: "Soil-Mates"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Packages
```{r}
library(tidyr)
library(tibble)
library(dplyr)
library(vegan)
library(tidyverse)
library(ggpubr)
library(ggtree)
library(vegan)
library(ape)
library(ggplot2)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(ggtext)
```

#### Importing data files at once
```{r}
temp <- list.files("./Data",pattern="*.csv", full.names = T) #creates an object that contains a string of file names with directory so they can be located
Dat <- list() # creates an empty vector used in the for loop
for (i in seq_along(temp)) { # creates a list of dataframes
  Dat[[i]] <- read.csv(
    file = temp[[i]]
  )
}
names <- c("En10", "En20", "Ar10", "Ba10", "Fu10", "Ar20", "Ba20", "Fu20", "En") 
Dat <- set_names(Dat, names) # sets the names for the data frames in Dat
```
#### Preprocessing before data is combined
```{r}
Dat$Fu10 <- Dat$Fu10 %>% #Removes extras from Fungi
  select(-Trophic.Mode,-Guild,-Confidence.Ranking)
```
```{r}
Dat$Fu20 <- Dat$Fu20 %>% #Removes extras from Fungi
  select(-Trophic.Mode,-Guild,-Confidence.Ranking) 
```
```{r}
Dat$Ba10 <- Dat$Ba10 %>% # Fixes capitalization
  rename(Cluster=(cluster)) 
```
```{r}
Dat$Ar10 <- Dat$Ar10 %>% # Fixes capitalization
  rename(Phyla=(phylum),
         Class=(class),
         Order=(order),
         Family=(family))
```
```{r}
Dat$En10 <- Dat$En10 %>% # Possibly unnecessary
  rename(Archaea=(Archaea_010cm),
         Bacteria=(Bacteria_010cm),
         Fungi=(Fungi_010cm))
```
```{r}
Dat$En20 <- Dat$En20 %>% # Makes so that when both Enviro tables are combined names match
  rename(Clearcut=(CLF),
         Salvage=(Salv_log),
         Organic.Carbon=(Organic.Carbon_prop),
         Exc.Magnesium=(Exc..Magnesium),
         pH.CaCl2.=(pH.Level..CaCl2.),
         Archaea=(Archaea_020cm),
         Bacteria=(Bacteria_020cm),
         Fungi=(Fungi_020cm)) %>%
  mutate(TSD= as.numeric(TSD)-1)
```
#### Combines datasets - reproducibility likely can be enhanced - this one uses the Importing all at once
```{r}
En<-bind_rows(Dat$En10,Dat$En20) #Combines Enviro tables into one dataset
D10<-bind_rows(Dat$Ar10,Dat$Ba10,Dat$Fu10) # Combines all for at depth 10 
D20<-bind_rows(Dat$Ar20,Dat$Ba20,Dat$Fu20) # Combines all microbes for for depth of 20
```
 
## Question 1
 
#### Begin diversity calculations by transposing data
```{r}
tD10 <- D10 #creates new object to be manipulated for diversity calculations
tD10 <- tD10 %>%
  select(-Cluster, -Sample, -Phyla, -Class, -Order, -Family)
tD10 <- t(tD10) #Transposes file
```

# calculates diversity for depth 0-10cm
```{r}
Div1 <- diversity(tD10, index = "shannon", MARGIN = 1, base= exp(1))
```

# transposes depth 20-30cm file
```{r}
tD20 <- D20
tD20 <- tD20 %>%
  select(-Cluster, -Sample, -Phyla, -Class, -Order, -Family)
tD20 <- t(tD20)
```

#calculates diveristy for depth of 20-30cm
```{r}
Div2 <- diversity(tD20, index = "shannon", MARGIN = 1, base= exp(1))
```

## T-test of diversity between depth
```{r}
t.test(Div1, Div2)
```

#### Creates a new environmental table and adds diversity data to it
```{r}
EnQ1 <- En
EnQ1$Div <- c(Div1, Div2)
```

#### Add proper names to a new column for disturbance History in a New environemental table using a for loop
```{r}
EnM <- EnQ1
for (i in 1:nrow(EnM)) { 
  if (EnM$TSD[i] == 167) {
  EnM$Disturbance_History[i] <- "1850F"
  } else if (EnM$TSD[i] == 78) {
  EnM$Disturbance_History[i] <- "1939F" 
  } else if ((EnM$No.Fires[i]== 2) && (EnM$TSD[i] == 34)) {
  EnM$Disturbance_History[i] <- "1939/1983F"
  } else if ((EnM$No.Fires[i]== 2) && (EnM$Salvage[i] == 0) && (EnM$Clearcut[i] == 0) && (EnM$TSD[i] == 8)) {
  EnM$Disturbance_History[i] <- "1939/2009F" 
  } else if (EnM$No.Fires[i]== 3) {
  EnM$Disturbance_History[i] <- "1939/1983/2009F" 
  } else if ((EnM$No.Fires[i]== 1) && (EnM$Clearcut[i] == 1) && (EnM$TSD[i] == 34)) {
  EnM$Disturbance_History[i] <- "1939F/1980CC"
  } else if ((EnM$No.Fires[i]== 1) && (EnM$Clearcut[i] == 1) && (EnM$TSD[i] == 8)) {
  EnM$Disturbance_History[i] <- "1939F/2009CC"
  } else if (EnM$Salvage[i] == 1) {
  EnM$Disturbance_History[i] <- "1939/2009F/2009SL"
  } else if ((EnM$No.Fires[i]== 2) && (EnM$Clearcut[i] == 1) && (EnM$TSD[i] == 8)) {
  EnM$Disturbance_History[i] <- "1939/1983F/2009CC"
  } 
}
```
### Stats for individual varibles
```{r}
two.way1 <- aov(Div ~ No.Fires + Depth, data = EnQ1)
summary(two.way1)
two.way2 <- aov(Div ~ TSD + Depth, data = EnQ1)
summary(two.way2)
two.way3 <- aov(Div ~ Clearcut + Depth, data = EnQ1)
summary(two.way3)
two.way4 <- aov(Div ~ Salvage + Depth, data = EnQ1)
summary(two.way4)
two.way5 <- aov(Div ~ Disturbance_History + Depth, data = EnM)
summary(two.way5)
```

#### base theme for boxplots
```{r}
theme_boxplot_one <- function (base_size =12, base_family = ""){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.text = element_text(colour = "black"),
      axis.title.x = element_text(size=12),
      axis.text.x = element_text(size=9),
      axis.title.y = element_text(size=12, angle=90),
      axis.text.y = element_text(size=9),
      axis.ticks = element_blank(),
      panel.background = element_rect(fill = "white"),
      panel.border = element_blank(),
      plot.title= element_text(face="bold", size=24)
    )
}
```

#Colour palettes for graphs
```{r}
Pal1 <- c("#7f4f24", "#718355")
Pal2 <- c("#e9972d", "#15616d")
fullpal <- c(Pal1, Pal2, "#732626", "#CB8282", "#BA82CB", "#82CBAA", "#C3CB82") 
```

#### Plot of diversity by depth
```{r}
EnM_mean <- EnM %>% 
  group_by(Depth) %>% 
  summarize(average = mean(Div)) %>%
  ungroup()
```

#### First plot comparing alpha diverisity between depths
```{r}
QP1 <- ggplot(EnM, aes(x=Depth, y=Div)) +
    geom_boxplot(aes(colour=I("black"), fill=Depth)) +  
    labs(y=NULL, 
         x="Depth of sample",
         fill="Disturbance History") + 
  scale_x_discrete(labels=c("0-10cm", "20-30cm")) +
        theme_boxplot_one() + 
  geom_point(data = EnM_mean, 
             mapping = aes(x= Depth, y = average),
             color="red") +
   geom_line(data = EnM_mean, 
            mapping = aes(x = Depth, y = average, group=1)) + 
  stat_compare_means(method = "t.test", label.x= 0.5) + 
  theme(legend.position = "none") +
  scale_fill_manual(values = Pal1)
QP1
```

#### Examing individual disturbance types effects
```{r}
plot1 <- ggplot(EnM, aes(x = as.factor(No.Fires), y = Div )) +
  stat_summary(aes(group = Depth, color = Depth, linetype = Depth), fun = mean, geom = "path", size =1) +
  stat_summary(aes(color = Depth, shape = Depth), fun = mean,
               geom = "errorbar", fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),  size =1, width=.25,
               position=position_dodge(width=0.03)) +
  stat_summary(aes(color = Depth, shape = Depth), geom = "point",fun.y = "mean",size = 4) +
  scale_shape_manual(name = "Group", values = c(18, 16)) +
    xlab ('Number of Fire') +
  ylab(NULL) +
    theme_boxplot_one() +
  guides(shape = FALSE, linetype = FALSE) +
  theme(text = element_text(size = 14)) +
    guides(colour = guide_legend(override.aes = list(shape = c(18, 16)), 'Depth')) + 
  theme(legend.position = "none") +
  scale_colour_manual(values = fullpal, labels = c("0-10cm", "20-30cm"))

plot2 <- ggplot(EnM, aes(x = as.factor(Clearcut), y = Div )) +
  stat_summary(aes(group = Depth, color = Depth, linetype = Depth), fun = mean, geom = "path", size =1) +
  stat_summary(aes(color = Depth, shape = Depth), fun = mean,
               geom = "errorbar", fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),  size =1, width=.25,
               position=position_dodge(width=0.03)) +
  stat_summary(aes(color = Depth, shape = Depth), geom = "point",fun.y = "mean",size = 4) +
  scale_shape_manual(name = "Group", values = c(18, 16)) +
    xlab ('Clearcut') +
  ylab(NULL) +
    theme_boxplot_one() +
  guides(shape = FALSE, linetype = FALSE) +
  theme(text = element_text(size = 14)) +
    guides(colour = guide_legend(override.aes = list(shape = c(18, 16)), 'Depth')) + 
  theme(legend.position = "none") +
  scale_colour_manual(values = fullpal, labels = c("0-10cm", "20-30cm"))

plot3 <- ggplot(EnM, aes(x = as.factor(Salvage), y = Div )) +
  stat_summary(aes(group = Depth, color = Depth, linetype = Depth), fun = mean, geom = "path", size =1) +
  stat_summary(aes(color = Depth, shape = Depth), fun = mean,
               geom = "errorbar", fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),  size =1, width=.25,
               position=position_dodge(width=0.03)) +
  stat_summary(aes(color = Depth, shape = Depth), geom = "point",fun.y = "mean",size = 4) +
  scale_shape_manual(name = "Group", values = c(18, 16)) +
    xlab ('Salvage Logging') +
   ylab(NULL) +
    theme_boxplot_one() +
  guides(shape = FALSE, linetype = FALSE) +
  theme(text = element_text(size = 14)) +
    guides(colour = guide_legend(override.aes = list(shape = c(18, 16)), 'Depth')) + 
  theme(legend.position = "none") +
  scale_colour_manual(values = fullpal, labels = c("0-10cm", "20-30cm"))

plot4 <- ggplot(EnM, aes(x = as.factor(TSD), y = Div )) +
  stat_summary(aes(group = Depth, color = Depth, linetype = Depth), fun = mean, geom = "path", size =1) +
  stat_summary(aes(color = Depth, shape = Depth), fun = mean,
               geom = "errorbar", fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),  size =1, width=.25, 
               position=position_dodge(width=0.03)) +
  stat_summary(aes(color = Depth, shape = Depth), geom = "point",fun.y = "mean",size = 4) +
  scale_shape_manual(name = "Group", values = c(18, 16)) +
    xlab ('Time Since Disturbance (Years)') +
  ylab(NULL) +
    theme_boxplot_one() +
  guides(shape = FALSE, linetype = FALSE) +
  theme(text = element_text(size = 14)) +
    guides(colour = guide_legend(override.aes = list(shape = c(18, 16)), 'Depth')) + 
  theme(legend.position = "none") + 
  scale_colour_manual(values = fullpal, labels = c("0-10cm", "20-30cm"))
```

####creates a dummy plot so that grid figure can have a common legend
```{r}
plotx <- ggplot(EnM, aes(x=Depth, y=Div)) + # Makes a plot with a legend
    geom_boxplot(aes(colour=I("black"), fill=Depth)) +
  theme_boxplot_one() + 
  scale_fill_manual(values = Pal1)
extract_legend <- function(my_ggp) { #Custom function to extract just legends
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
# Apply user-defined function to extract legend for us in a mutliplot
shared_legend <- extract_legend(plotx)
```

#### creates a multiplot with all disturbance factor plots
```{r}
PlotB<- grid.arrange( arrangeGrob(plot1, plot2, plot3, plot4, ncol=2), 
                  shared_legend, nrow = 2, heights = c(5, 1))
             
```

#### Codes means based off of Disturbance_Histories
```{r}
EnM_mean2 <- EnM %>% 
  group_by(Disturbance_History) %>% 
  summarize(average = mean(Div)) %>%
  ungroup()
```

#### Diversity by disturbance histories
```{r}
QP6<- ggplot(EnM, aes(x=reorder(Disturbance_History,-TSD), y=Div)) +
    geom_boxplot(aes(colour=I("black"), fill=Disturbance_History)) +  
    labs(y=NULL, 
         x="Disturbance History",
         fill="Disturbance History") + 
   geom_point(data = EnM_mean2, 
             mapping = aes(x= Disturbance_History, y = average),
             color="red") +
   geom_line(data = EnM_mean2, 
            mapping = aes(x = Disturbance_History, y = average, group=1)) + 
  theme_boxplot_one() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 15, vjust = 1, hjust=1)) +
  scale_fill_manual(values = fullpal)
QP6
```

#### Creates a final multi plot
```{r}
PlotFin<-grid.arrange(QP1, PlotB, QP6, ncol=1, nrow=3, left = text_grob("Shannon's Diversity", size=20, rot=90))
```

#### Creates and saves a .pdf of the final plot that is properly sized
```{r}
pdf("Question1Plots.pdf") 

ggsave( "Question1Plots.pdf", plot=PlotFin, width =30, height = 30, units = "cm")
```

# Question 2. Are different soil chemical compositions associating with destructive forces?

##Introduction

## The analysis

Quick overview to check our dataframe
```{r}
summary(En)
```


Destructive forces :
No.Fires - number of fires either 1, 2, 3 or 4
Clearcut- is it clearcut yes = 1, no = 0
Salvage - was it salvage logged - yes = 1, no = 0
TSD is time since disturbance


Chemical info:
colwell is a way of measuring availibility of that element
DTPA is Diethylenetriaminepentaacetic acid which extracts trace elements


Here I am changing destructive forces to characters, for boxplotting

```{r}
En$no_fires <- as.character(En$No.Fires)
```

```{r}
En$clearcut <- as.character(En$Clearcut)
```

```{r}
En$Salvage <- as.character(En$Salvage)
```


#Exploratory boxplots, looking at any that could be of interest

```{r}
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(Ammonium.Nitrogen), fill = clearcut))
```

```{r}
#No difference
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = clearcut))
```

```{r}
#No fires has higher nitrogen in nitrate form compared to all number of fires greater than 2
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut))
```

```{r}
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(Phosphorus.Colwell), fill = clearcut))
```

```{r}
#three fires more potassium 
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell,fill = clearcut))
```

```{r}
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Copper), fill = clearcut))
```

```{r}
#lower iron 3 fires? 
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = clearcut))
```

```{r}
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Manganese), fill = clearcut))
```

```{r}
#higher magnesium with increasing number of fires 
ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(Exc.Magnesium), fill = clearcut))
```

#Picking the best graphs from the exploring, and making them look publication ready
## ADDING geom_jitter to see raw data on boxplots

```{r}
a <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the sitce 
clear cut?")
a
```

```{r}
b <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = clearcut),outlier.shape = NA)+
  geom_jitter(aes(x = no_fires,y = Potassium.Colwell, fill = clearcut), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
b
```


```{r}
c <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = clearcut), outlier.shape = NA)+
   geom_jitter(aes(x = no_fires, y = log10(DTPA.Iron)), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
clear cut?")
c
```
 
```{r}
d <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = clearcut), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Organic.Carbon, fill = clearcut), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
d
```

#SALVAGE LOGGING

```{r}
e <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = Salvage), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Organic.Carbon, fill = Salvage), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
salvage logged??")
e
```

```{r}
f <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = Salvage), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = log10(DTPA.Iron), fill = Salvage),width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
salvage logged?")
f
```

```{r}
g <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = Salvage), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Potassium.Colwell, fill = Salvage), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")
g
```

```{r}
h <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = Salvage),outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Nitrate.Nitrogen, fill = Salvage), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")
h
```


#if we are stacking we can lose things that are the same between graphs

```{r}
a <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
b <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = clearcut),outlier.shape = NA)+
  geom_jitter(aes(x = no_fires,y = Potassium.Colwell, fill = clearcut), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
c <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = clearcut), outlier.shape = NA)+
   geom_jitter(aes(x = no_fires, y = log10(DTPA.Iron)), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
clear cut?")
d <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = clearcut), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Organic.Carbon, fill = clearcut), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal1)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
h <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = Salvage),outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Nitrate.Nitrogen, fill = Salvage), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")

g <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = Salvage), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Potassium.Colwell, fill = Salvage), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")
f <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = Salvage), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = log10(DTPA.Iron), fill = Salvage),width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
salvage logged?")
e <- ggplot(En)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = Salvage), outlier.shape = NA)+
  geom_jitter(aes(x = no_fires, y = Organic.Carbon, fill = Salvage), width=0.1, shape=21, color="black")+
  scale_fill_manual(labels = c('No','Yes'), values = Pal2)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
salvage logged??")
```

## Stacking
```{r}
grid.arrange(a, b, c, d,
             ncol = 1, nrow = 4)
```

## QUESTION 3: Can we associate changes in microbial communities with soil composition?

- We are interested in how community level diversity changes when considering destructive events and soil compositions
- Primarily, we are looking for how tree branche and points cluster

### Preprocessing

Searching for the OTU table .csv files
```{r}
# We initialize the variable temp
temp =0

# We fill temp with a list of the files which have the string pattern. 
  # We use a wild card (*) in order to allow generality (depths and archaea, bacteria and fungi)
temp <- list.files("./Data",pattern="Dryad_Microbial_OTU+.*csv", full.names = T)

# Initializing Dat which will hold all the dataframes
Dat <- c()
```

Retrieving the dataframes using our temp list
```{r}
# Loops through temp, reads .csv into Dat.
for (i in seq_along(temp)) {
  Dat[[i]] <- read.csv(file = temp[[i]])
  # Each item in Dat is pulled out as it's own dataframe. We use sub and assign to make unique names for these dataframes
  data.frame(assign(sub(".csv", "",sub("./Data/Dryad_Microbial_OTU_Table*", "", temp[i])), data.frame(Dat[i])))

}
```

The fungi OTU tables have columns which group the Families by known function. For our purposes, we won't be needing them, therefore we dropped them.
```{r}
s_10_Fungi <- subset(s_10_Fungi, select = -c(Trophic.Mode, Guild, Confidence.Ranking))
s_20_Fungi <- subset(s_20_Fungi, select = -c(Trophic.Mode, Guild, Confidence.Ranking))
```

In order to allow the merging of different types of microbes, we need to standardize the column names. We created the custom function,upper, in order to capitalize all the column names. 
```{r}
# Input: The dataframe
upper <- function(df) {
  
  # Loops through all the column names
for( i in colnames(df)){
  # Uses the toupper function to capitalize the column name we're looping through (overwrites)
    colnames(df)[which(colnames(df)==i)] = toupper(i)
}
  # Returns the dataframe
    return (df)
}
```

Calling uppers function for all our OTU tables

```{r}
s_10_Archaea = upper(s_10_Archaea)
s_20_Archaea = upper(s_20_Archaea)
s_10_Bacteria = upper(s_10_Bacteria)
s_20_Bacteria = upper(s_20_Bacteria)
s_10_Fungi = upper(s_10_Fungi)
s_20_Fungi = upper(s_20_Fungi)
```

Bowd, et al, (2020) preprocessed the data using UNITE and Greengenes, up to the Family level. This resulted multiple OTU counts which are non-unique, and can be agglomerated. In order to do this, we pasted the Phyla, class, order and Family names for each row. 

```{r}
# Interestingly, only the s_10_Archaea table has a unique naming system for Phylum
colnames(s_10_Archaea)[colnames(s_10_Archaea) == 'PHYLUM'] <- 'PHYLA'

s_10_Archaea$Phylogeny <- paste0(s_10_Archaea$PHYLA, s_10_Archaea$CLASS, s_10_Archaea$ORDER, s_10_Archaea$FAMILY, sep = "")
s_20_Archaea$Phylogeny <- paste0(s_20_Archaea$PHYLA, s_20_Archaea$CLASS, s_20_Archaea$ORDER, s_20_Archaea$FAMILY, sep = "")

s_10_Bacteria$Phylogeny <- paste0(s_10_Bacteria$PHYLA, s_10_Bacteria$CLASS, s_10_Bacteria$ORDER, s_10_Bacteria$FAMILY, sep = "")
s_20_Bacteria$Phylogeny <- paste0(s_20_Bacteria$PHYLA, s_20_Bacteria$CLASS, s_20_Bacteria$ORDER, s_20_Bacteria$FAMILY, sep = "")

s_10_Fungi$Phylogeny <- paste0(s_10_Fungi$PHYLA, s_10_Fungi$CLASS, s_10_Fungi$ORDER, s_10_Fungi$FAMILY, sep = "")
s_20_Fungi$Phylogeny <- paste0(s_20_Fungi$PHYLA, s_20_Fungi$CLASS, s_20_Fungi$ORDER, s_20_Fungi$FAMILY, sep = "")
```

In preparation for merging different depths, we need unique columns names for each site. We created a custom function, changenames, which uses sub to add the depth, one of the inputs.

```{r}
# Input: dataframe and depth
changenames <- function(df, depth) {
  # Loops through all the columns
for ( col in 1:ncol(df)){
  # Adds the depth before the X character
    colnames(df)[col] <-  sub("X", depth, colnames(df)[col])
    
}
  # Returns the dataframe
  return (df)
}
```

Calling changenames function for all our OTU tables

```{r}
s_10_Archaea = changenames(s_10_Archaea, 'depth10_')
s_20_Archaea = changenames(s_20_Archaea, 'depth20_')

s_10_Bacteria = changenames(s_10_Bacteria, 'depth10_')
s_20_Bacteria = changenames(s_20_Bacteria, 'depth20_')

s_10_Fungi = changenames(s_10_Fungi, 'depth10_')
s_20_Fungi = changenames(s_20_Fungi, 'depth20_')
```

As mentioned, there are a number of non-unique phylogenies, likely because of varying resolution upon using the reference. Therefore, we created a custom function, groupby_ing, to groupby unique phylogenies and to sum across the units.

```{r}
# Input: dataframe
groupby_ing <- function(df) {
  
  # Uses dplyr functions
  df <- df %>% 
    # Groups by "complete" Phylogeny 
      group_by(Phylogeny) %>%
    # Row-wise sum of units
  summarise_if(is.numeric, sum) %>% 
    # Removal of the Cluster column since we won't need it
  select (-CLUSTER)
  # Returns the dataframe
  return (df)

}
```

Calling groupby_ing function for all our OTU tables

```{r}
s_10_Archaea_agglom = groupby_ing(s_10_Archaea)
s_20_Archaea_agglom = groupby_ing(s_20_Archaea)

s_10_Bacteria_agglom = groupby_ing(s_10_Bacteria)
s_20_Bacteria_agglom = groupby_ing(s_20_Bacteria)

s_10_Fungi_agglom = groupby_ing(s_10_Fungi)
s_20_Fungi_agglom = groupby_ing(s_20_Fungi)
```

Prep for a Validation check: How many different phylogenies are present?

- Among each type of microbe (archaea, bacteria and fungi), we know the microbes found may be different at different depths. 
- Once we merge, we'll check if the number of unique phylogenies match


```{r}
# For archaea
dim(s_10_Archaea_agglom)
dim(s_20_Archaea_agglom)
# find the number of unique phylogenies, we should expect 19 upon merging
length(unique(c(s_10_Archaea_agglom$Phylogeny, s_20_Archaea_agglom$Phylogeny)))

# For bacteria
dim(s_10_Bacteria_agglom)
dim(s_20_Bacteria_agglom)
# find the number of unique phylogenies, we should expect 419 upon merging
length(unique(c(s_10_Bacteria_agglom$Phylogeny, s_20_Bacteria_agglom$Phylogeny)))

# For fungi
dim(s_10_Fungi_agglom)
dim(s_20_Fungi_agglom)
# find the number of unique phylogenies, we should expect 321 upon merging
length(unique(c(s_10_Fungi_agglom$Phylogeny, s_20_Fungi_agglom$Phylogeny)))
```

In order to conduct diversity analysis (Phylogenetic trees and Bray-Curtis), we'll need to transpose all our OTU tables, where phylogenies are the column. Therefore, we created a custom function, transposing, to transpose the OTU tables. 

```{r}
# Input: Dataframe
transposing <- function(df) {
  
  # Transposing the dataframe on the Phylogeny column, site IDs are now under the column name, Phylogeny.
df = transpose(setDT(df), make.names = 'Phylogeny', keep.names = 'Phylogeny')
  
  # Returns the dataframe
  return (df)
}
```

Calling groupby_ing function for all our agglomerated OTU tables

```{r}
s_10_Archaea_agglom_t = transposing(s_10_Archaea_agglom)
s_20_Archaea_agglom_t = transposing(s_20_Archaea_agglom)

s_10_Bacteria_agglom_t = transposing(s_10_Bacteria_agglom)
s_20_Bacteria_agglom_t = transposing(s_20_Bacteria_agglom)

s_10_Fungi_agglom_t = transposing(s_10_Fungi_agglom)
s_20_Fungi_agglom_t = transposing(s_20_Fungi_agglom)
```

In order to understand how diversity differs among microbe type, we will agglomerate the OTUs from each depth. In addition, we'll make an "all" table with all three microbe types at both depths. 

- We will also check the unique number of phylogenies, and compare with the above values. 

```{r}
# Row binding both 
arc = rbind(s_10_Archaea_agglom_t, s_20_Archaea_agglom_t,fill=TRUE)
# checking unique number, we are expecting 19
ncol(arc) -1

bac = rbind(s_10_Bacteria_agglom_t, s_20_Bacteria_agglom_t,fill=TRUE)
# checking unique number, we are expecting 419
ncol(bac) -1

fun = rbind(s_10_Fungi_agglom_t, s_20_Fungi_agglom_t,fill=TRUE)
# checking unique number, we are expecting 321
ncol(fun) -1

# Merging all together
all = list(arc,bac,fun ) %>% 
  reduce(full_join, by = "Phylogeny")
```

Adding column for depth with a function

```{r}
add_depth <- function(df) {
  
  df <- df %>% 
      mutate(depth = ifelse(grepl("depth10_", Phylogeny), '10',"20")) 
  
  return (df)

}
```

Calling add_depth function for the agglomerated OTUs

```{r}
arc = add_depth(arc)
bac = add_depth(bac)
fun = add_depth(fun)

all = add_depth(all)
```

In order for us to compare features with beta diversity, we'll create a "df" dataframe which we will merge with the environmental data. We'll pull from the bacteria dataframe

```{r}
df = bac[, c("Phylogeny", "depth")]
```

We are pulling the environmental data and making each site match the format used in the OTU tables

```{r}
# Using dplyr and an ifelse, baseed on the depth, we'll paste depth and site ID
En = En %>% 
  mutate(Phylogeny = ifelse(grepl("010cm", Depth), paste0('depth10_',Sample), paste0('depth20_',Sample))) 
```

Finally merging the df created with the OTU tables and the environmental table. Now we have a "describe" dataframe which holds all the soil composition data. 

```{r}
# We merged on the column, "Phylogeny" which is the sample ID
describe = merge(df, En, by = 'Phylogeny')
```

When conducting our diversity analysis, we will use the Sample IDs (in the Phylogeny column) as the identifier. Therefore, we will convert this column into the row names, instead of the index. 

```{r}
# Input: Dataframe
rowname <- function(df) {
  # We'll use dplyr to remove the current rownames and replac with the Phylogenny column
  df = df %>% 
  remove_rownames %>% 
  column_to_rownames(var="Phylogeny")
  
  # Returns the dataframe
  return (df)

}
```

Calling rowname function for the agglomerated OTUs

```{r}
arc = rowname(arc)
bac = rowname(bac)
fun = rowname(fun)

all = rowname(all)
```

Upon merging, all rows, not found, recieve a NA by default. The vegdist function we will eventually use, has difficulties processing NAs, therefore we will convert these cells into 0s with the custom function, nonan.

```{r}
# Input: dataframe
nonan <- function(df) {
  # Uses dplyr
df = df %>%
  # For all numeric columns (ie. those with units), convert NAs into 0s
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), 0, .)))
  # Return the dataframe
  return (df)

}
```

Calling nonnan function for all the agglomerated OTU tables

```{r}
arc_nonan = nonan(arc)
bac_nonan = nonan(bac)
fun_nonan = nonan(fun)
all_nonan = nonan(all)
```

We are interested in assessing beta-diversity or the measurement of diversity when comparing communities. We will use the vegdist function to calculate Bray-Curtis and use the nj function to create Neighbour-Joining trees. This will be efficiently calculated using the custom function, OTU_bray.


```{r}
# Input: Dataframe
OTU_bray <- function(df) {
  
  # Calculating Bray-Curtis using ONLY the microbes (we subset the dataframe to drop depth)
  OTU_dist<- vegdist(subset(df, select = -depth),method="bray",binary=F) 
  # We apply the NJ function
  OTU_tree2<-nj(OTU_dist)
  
  # We return a two item list: the Bray-Curtis distances and the tree
   return(list(OTU_dist,OTU_tree2))
}
```

Calling OTU_bray function for all the agglomerated OTU tables. We will also subset the OTU tables by depth, as we hypothesize depth is heavily influential to microbial communities, and the calculation for "differences" will be different via subsetting. 

```{r}
arc_nonan_bray = OTU_bray(arc_nonan)
bac_nonan_bray = OTU_bray(bac_nonan)
fun_nonan_bray = OTU_bray(fun_nonan)

all_nonan_bray = OTU_bray(all_nonan)

# By depth

# bacteria
bac_nonan_bray_depth10 = OTU_bray(bac_nonan[bac_nonan$depth ==10,])
bac_nonan_bray_depth20 = OTU_bray(bac_nonan[bac_nonan$depth ==20,])

# for all
all_nonan_bray_depth10 = OTU_bray(all_nonan[all_nonan$depth ==10,])
all_nonan_bray_depth20 = OTU_bray(all_nonan[all_nonan$depth ==20,])
```

Could make this a loop... across diff features ** ONly did it for arc_nonan_bray, repeat for other two df...

Using the above, we will assess clustering visually using the tree calculated above (the 2nd object in the list from OTU_bray) We will do this efficiently by creating a list of the categorical values (ie. related to destructive forces) and continuous values (ie. mostly soil composition measurements)  then colouring each tip. 

```{r}
# Categorical values
features = c('depth', 'No.Fires', 'Clearcut', 'Salvage')

# Creating a tree using each feature
for (i in features) {
  print(ggtree(bac_nonan_bray_depth20[[2]]) %<+% 
  describe + 
  geom_tippoint(aes(color=as.factor(get(i)) ))+
  guides(color=guide_legend(title=i)))
}

# Continuous features
cont = c('TSD', 'Ammonium.Nitrogen', 'Nitrate.Nitrogen', 'Phosphorus.Colwell', 'Potassium.Colwell', 'Organic.Carbon', 'DTPA.Copper', 'DTPA.Iron','DTPA.Manganese',  'Exc.Magnesium','Slope', 'Elevation', 'pH.CaCl2.')

for (i in cont) {
  print(ggtree(arc_nonan_bray[[2]]) %<+% 
  describe + 
  geom_tippoint(aes(color=get(i) ))+
  guides(color=guide_legend(title=i)))
}
```

From this:
- Few of these features createdistinct branches
- Few sites who underwent salvage cutting


Next, we will prepare the NMDS cluters, however we want to keep the unique sample IDs. For simplicity, we'll check if we can use only one of the OTU tables, by checking if the order of the Phylogeny column is the same. 

```{r}
# The identical function checks to see if all the values and their order match. We will try this on all the relevant combinations. 
identical(row.names(arc), row.names(bac) )
identical(row.names(arc), row.names(fun) )
identical(row.names(arc), row.names(all) )
```

Since all the values are true, we will only use the rownames from the arc dataframe.

Calculating NMDS\
We are interested in studying how the microbial communities cluster, therefore we will apply non-metric multidimensional scaling (NMDS) which reduces the number of dimensions. We will apply this to all our OTU tables using the custom function, NMDSdat_make. We played around with the number of runs (trymax) in order to reach a solution for each table. 

```{r}
# Input: Dataframe and the dataframe we'll use for rownames (arc)
NMDSdat_make <- function(df, rowwing) {
  # We'll set a seed to allow for reproducibility
  set.seed(13)
  
  # At 20 runs, fails to converge, 
  #NMDSdat<-metaMDS(df,k=2) # k = 2 dimensions
  # All fails to converge with 100 runs, so we'll increase to 200
  #NMDSdat<-metaMDS(df,k=2, trymax = 100) # k = 2 dimensions
  
  # We found at 500 runs, all the tables reach a solution
  NMDSdat<-metaMDS(df,k=2, trymax = 500) # k = 2 dimensions

  # We create a dataframe to hold all the NMDS values
  PDat<-data.frame(NMDS1=NMDSdat$points[,1],
                 NMDS2=NMDSdat$points[,2],
                 Phylogeny=row.names(rowwing))
  # Return the PDat dataframe
return(PDat)
}
```


Calling NMDSdat_make for all the agglomerated OTU tables

```{r}
NMDSdat_arc = NMDSdat_make(arc_nonan_bray[[1]], arc)
NMDSdat_bac = NMDSdat_make(bac_nonan_bray[[1]], arc)
NMDSdat_fun = NMDSdat_make(fun_nonan_bray[[1]], arc)

NMDSdat_all = NMDSdat_make(all_nonan_bray[[1]], arc)

# By depth

# For bacteria
NMDSdat_bac_depth10 = NMDSdat_make(bac_nonan_bray_depth10[[1]], arc[1:80,])
NMDSdat_bac_depth20 = NMDSdat_make(bac_nonan_bray_depth20[[1]], arc[81:160,])


# For all
NMDSdat_all_depth10 = NMDSdat_make(all_nonan_bray_depth10[[1]], arc[1:80,])
NMDSdat_all_depth20 = NMDSdat_make(all_nonan_bray_depth20[[1]], arc[81:160,])
```

In order to aadd soil composition data, we will merge the NMDS data with the "describe" dataframe, based on the Sample ID. 

```{r}
NMDSdat_arc = merge(NMDSdat_arc, describe, by = 'Phylogeny')
NMDSdat_bac = merge(NMDSdat_bac, describe, by = 'Phylogeny')
NMDSdat_fun = merge(NMDSdat_fun, describe, by = 'Phylogeny')

NMDSdat_all = merge(NMDSdat_all, describe, by = 'Phylogeny')

NMDSdat_bac_depth10 = merge(NMDSdat_bac_depth10, describe, by = 'Phylogeny')
NMDSdat_bac_depth20 = merge(NMDSdat_bac_depth20, describe, by = 'Phylogeny')

NMDSdat_all_depth10 = merge(NMDSdat_all_depth10, describe, by = 'Phylogeny')
NMDSdat_all_depth20 = merge(NMDSdat_all_depth20, describe, by = 'Phylogeny')
```

Similar to the above, we made a loop to dissect and cluster by categorical features, we'll apply this to all the OTU tables with the custom function, viz_cats. 

```{r}
features = c('depth', 'No.Fires', 'Clearcut', 'Salvage')

# Input: dataframe
viz_cats <- function(df) {
  
  # Looping through the features list
for (i in features) {
  # Make a clustering plot each feature
  print(qplot(x=NMDS1,y=NMDS2,color=as.factor(get(i)), alpha= I(0.6), data= df)+
  stat_ellipse() + 
  theme_bw()+
   guides(color=guide_legend(title=i))+
    ggtitle(deparse(substitute(df))))
}
}
```

Calling viz_cats for all NMDS dataframes. 

```{r}
viz_cats(NMDSdat_arc)
viz_cats(NMDSdat_bac)
viz_cats(NMDSdat_fun)
viz_cats(NMDSdat_all)

viz_cats(NMDSdat_bac_depth10)
viz_cats(NMDSdat_bac_depth20)

viz_cats(NMDSdat_all_depth10)
viz_cats(NMDSdat_all_depth20)
```

From the above, we determined:
- Depths 0-10cm versus 20-30cm, makes a big difference when comparing clustering (This is in part our rational for separating Bray-Curtis calculations by depth). 
  - This trend is especially obvious when observing bacterial communities
    - We will therefore focus on these NMDS dataframes
  - This trend seems to correlate with the amount of organic carbon (not shown)
- We additionally noticed, among both depths, Clearcut 0 and 1 (no vs yes) has some distinct clustering
- We found the number of fires does not result in unique clustering

We'll repeat the above with continuous variables to determine if clustering is linked to a soil health indicator. 

```{r}
# List of continuous variables
cont = c('Soil_fertility','TSD', 'Ammonium.Nitrogen', 'Nitrate.Nitrogen', 'Phosphorus.Colwell', 'Potassium.Colwell', 'Organic.Carbon', 'DTPA.Copper', 'DTPA.Iron','DTPA.Manganese',  'Exc.Magnesium','Slope', 'Elevation', 'pH.CaCl2.')

# Same function, but with this list (no as.factor function required)
viz_cont <- function(df) {
  
for (i in cont) {
  print(ggplot(data = df, aes(NMDS1, NMDS2 )) + 
  geom_point(aes(color=get(i), shape=as.factor(Clearcut), size=2, alpha=I(0.7) )) +
  labs(color=i)+
    ggtitle(deparse(substitute(df)))+
  theme_bw())
}
}
```

Calling viz_cont for the two NMDS dataframes of interest. 

```{r}
viz_cont(NMDSdat_bac_depth10)
viz_cont(NMDSdat_bac_depth20)
```

From above, we can see:
- High vs low pH seems to correlate with the distinct clusters from Clearcutting.
  - Additionally, this effect is more present among 0-10cm depth 
- Nitrogen potentially clusters distinctly, but only loosely


As we believe pH is related to Clearcutting, we will refine our final figures for this finding. 

```{r}
# At depth 0-10cm

a=  ggplot(data = NMDSdat_bac_depth10, aes(NMDS1, NMDS2 )) + 
  geom_point(aes(color=pH.CaCl2., shape=as.factor(Clearcut), size=2, alpha=I(0.7)  )) +
          stat_ellipse(aes(group=as.factor(Clearcut))) +
  xlim(c(-0.4,0.35))+
  ylim(c(-0.25,0.2))+
  theme_bw()+
  theme(legend.position="none", text = element_text(size=20))+
  # Customize a colour gradient, similar to a litmus test
    scale_colour_gradient(
  low = "#e32636",
  high = "#5d8aa8")+
  # Adding annotations for clearcut/ not clear cut and depth
  annotate("text", x = -0.15, y = -0.15, label = "Clearcut", size=5)+
annotate("text", x = 0.3, y = -0.1, label = "No Clearcut", size=5)+
  annotate("text", x = 0.2, y = 0.2, label = "0-10cm depth", fontface = "bold", size=7)

# Repeat for 20-30cm depth
b=ggplot(data = NMDSdat_bac_depth20, aes(NMDS1, NMDS2 )) + 
  geom_point(aes(color=pH.CaCl2., shape=as.factor(Clearcut), size=4, alpha=I(0.7) )) +
            stat_ellipse(aes(group=as.factor(Clearcut))) +
  xlim(c(-0.4,0.35))+
  ylim(c(-0.25,0.2))+
  theme_bw()+
      scale_colour_gradient(
  low = "#e32636",
  high = "#5d8aa8")+
    labs(shape="Clearcut", colour="pH")+
    scale_shape_discrete(labels = c("No", "Yes"))+
  scale_size(guide = 'none')+
annotate("text", x = 0.2, y = -0.1, label = "Clearcut", size=5)+
annotate("text", x = 0.2, y = 0.1, label = "No Clearcut", size=5)+
    annotate("text", x = 0.2, y = 0.2, label = "20-30cm depth", fontface = "bold", size=7)+
  theme(legend.position="bottom",text = element_text(size=20))
```

Making subplots for each depth

```{r}
c = plot_grid(a, b, nrow = 2,labels = "AUTO",   rel_heights = c(4,5))

# Addding a caption
ggdraw(add_sub(c, size =20,
               'Figure 4. Non-metric multidimensional scaling (NMDS) ordination plot \n depicting clusters and shape which distinguishes \n sites which have undergone clearcutting. \n Color gradient represents pH measured in CaCl2. \n (A) 0-10cm depth samples. (B) 20-30cm depth samples')) 

# Saving the final figure as a .png
ggsave("./Question 1/fig4.png", width =24, height = 40, units = "cm", limitsize = FALSE)
```