---
title: "432analysis"
author: "Logan Wisteard"
date: '2022-04-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 2. Are different soil chemical compositions associating with destructive forces?

##Introduction

## The analysis

Load packages
```{r}
library(dplyr)
library(ggplot2)
library(ggtree)
library(vegan)
library(ape)
library(gridExtra)
```

We are calling the dataframe "En.csv" "en" for our analysis and retrieving it from the Data folder.
```{r}
en <- read.csv("../Data/En.csv")
```

Quick overview to check our dataframe
```{r}
summary(en)
```


Destructive forces :
No.Fires - number of fires either 1, 2, 3 or 4
Clearcut- is it clearcut yes = 1, no = 0
Salvage - was it salvage logged - yes = 1, no = 0
TSD is time since disturbance


Chemical info:
colwell is a way of measuring availibility of that element
DTPA is Diethylenetriaminepentaacetic acid which extracts trace elements


Here I am changing destructive forces to characters, for boxplotting

```{r}
en$no_fires <- as.character(en$No.Fires)
```

```{r}
en$clearcut <- as.character(en$Clearcut)
```

```{r}
en$Salvage <- as.character(en$Salvage)
```


#Exploratory boxplots, looking at any that could be of interest

```{r}
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(Ammonium.Nitrogen), fill = clearcut))
```

```{r}
#No difference
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = clearcut))
```

```{r}
#No fires has higher nitrogen in nitrate form compared to all number of fires greater than 2
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut))
```

```{r}
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(Phosphorus.Colwell), fill = clearcut))
```

```{r}
#three fires more potassium 
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell,fill = clearcut))
```

```{r}
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Copper), fill = clearcut))
```

```{r}
#lower iron 3 fires? 
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = clearcut))
```

```{r}
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Manganese), fill = clearcut))
```

```{r}
#higher magnesium with increasing number of fires 
ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(Exc.Magnesium), fill = clearcut))
```

#Colour palettes for graphs

```{r}
clearcutpal <- c("#7f4f24", "#718355")
```

```{r}
salvagepal <- c("#e9972d", "#15616d")
```

#base theme for boxplots

```{r}
theme_boxplot_one <- function (base_size =12, base_family = ""){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.text = element_text(colour = "black"),
      axis.title.x = element_text(size=12),
      axis.text.x = element_text(size=9),
      axis.title.y = element_text(size=12, angle=90),
      axis.text.y = element_text(size=9),
      axis.ticks = element_blank(),
      panel.background = element_rect(fill = "white"),
      panel.border = element_blank(),
      plot.title= element_text(face="bold", size=24)
    )
}
```

#Picking the best graphs from the exploring, and making them look publication ready

## ADDING outlier.alpha =0.5 to see if outlier may be one point or more than one point

```{r}
a <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut), outlier.alpha =0.5)+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
a
```

```{r}
b <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = clearcut))+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
b
```


```{r}
c <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = clearcut))+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
clear cut?")
c
```

```{r}
d <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = clearcut))+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
d
```

#SALVAGE LOGGING

```{r}
e <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = Salvage))+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
salvage logged??")
e
```

```{r}
f <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = Salvage))+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
salvage logged?")
f
```

```{r}
g <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = Salvage))+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")
g
```

## ADDING outlier.alpha =0.5 to see if outlier may be one point or more than one point
```{r}
h <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = Salvage), outlier.alpha =0.5)+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")
h
```


#if we are stacking we can lose things that are the same between graphs

```{r}
a <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = clearcut))+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
b <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = clearcut))+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
c <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = clearcut))+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
clear cut?")
d <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = clearcut))+
  scale_fill_manual(labels = c('No','Yes'), values = clearcutpal)+
  theme_boxplot_one()+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
clear cut?")
h <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Nitrate.Nitrogen, fill = Salvage))+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Nitrate form of Nitrogen (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")
g <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Potassium.Colwell, fill = Salvage))+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Potassium (ppm)\n")+
  labs(fill = "Was the site 
salvage logged?")
f <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = log10(DTPA.Iron), fill = Salvage))+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
   theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = expression(paste(log[10]," (Iron ppm)")))+
  labs(fill = "Was the site 
salvage logged?")
e <- ggplot(en)+
  geom_boxplot(aes(x = no_fires, y = Organic.Carbon, fill = Salvage))+
  scale_fill_manual(labels = c('No','Yes'), values = salvagepal)+
  theme_boxplot_one()+
  theme(legend.position = "none")+
  scale_x_discrete(name = "\nNumber of forest fires")+
  scale_y_continuous(name = "Organic Carbon (ppm)\n")+
  labs(fill = "Was the site 
salvage logged??")
```

#Stacking

```{r}
grid.arrange(a, b, c, d,
             ncol = 1, nrow = 4)
```
